Map{R}educe in {MPI} for Large-scale Graph Algorithms

### 2010-10-29 paper/results.tex
Updated with RMAT-20 results; still waiting for RMAT-28 runs.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@648 3fc257bb-8550-0410-81b3-a5c103da83f7

A more detailed analysis of the performance of the MR-MPI PageRank is shown
below.  We perform the analysis using RMAT-20, which is small enough to
fit in memory and thus does not need out-of-core operations, and RMAT-28,
which requires out-of-core operation in MR-MPI.  Table~\ref{table:prdetail}
shows the amount of time spent in each step of Figure~\ref{fig:pr2} for 
the MR-MPI implementation and the Trilinos implementation.

\begin{table}[htb]
\begin{center}
\begin{tabular}{|l|r|r|r|r|}
\hline 
 & \multicolumn{2}{|c|}{RMAT-20} & \multicolumn{2}{|c|}{RMAT-28} \\
 & MR-MPI & Trilinos & MR-MPI & Trilinos  \\
\hline
Step 1 (Compute adjustment) & & & &  \\
Step 2 (Multiply $A^T x$) & & & &  \\
Step 3 (Apply adjustment) & & & &  \\
Step 4 (Scale $y$)  & & & & \\
Step 5 (Compute residual)  & & & & \\
\hline
\end{tabular}
\caption{Detailed execution times in each stage of the PageRank algorithm
described in Figure~\ref{fig:pr2} using MR-MPI and Trilinos.}
\label{table:prdetail}
\end{center}
\end{table}

In Tables~\ref{table:prdetail2},
we break down the operation times further for the most expensive 
step of the PageRank computation, Step 2.
In the table, we divide the collate operation into its
two parts:  an aggregate that involves parallel communication and a convert
that gathers local key-value pairs with matching keys into a key-multivalue
pair.

\begin{table}[htb]
\begin{center}
\begin{tabular}{|l|r|r|}
\hline 
 & RMAT-20 & RMAT-28 \\
\hline
Copy $x$ to $y$ & &  \\
Add $A$ to $y$ & &  \\
Convert $y$ & &  \\
Reduce $y$ & & \\
Aggregate $y$ (first part of collate) & & \\
Convert $y$ (second part of collate) & & \\
Reduce $y$ & & \\
\hline
\end{tabular}
\caption{Detailed execution times of MR-MPI in Step 2 of 
Figure~\ref{fig:pr2}.}
\label{table:prdetial2}
\end{center}
\end{table}

Similarly, in Table~\ref{table:prdetail5}, we show the execution times 
required for each operation in Step 5 of Figure~\ref{fig:pr2}, the 
residual computation.

\begin{table}[htb]
\begin{center}
\begin{tabular}{|l|r|r|}
\hline 
 & RMAT-20 & RMAT-28 \\
\hline
Add $y$ to $x$ & &  \\
Convert $x$ & &  \\
Reduce $x$ & & \\
MPI\_Allreduce & & \\
\hline
\end{tabular}
\caption{Detailed execution times of MR-MPI in Step 5 of 
Figure~\ref{fig:pr2}.}
\label{table:prdetial2}
\end{center}
\end{table}
%This result is
%due to two factors: ({\it i}) a higher volume of communication in the
%MapReduce implementation (which comunicates one datum per graph edge
%in each iteration), compared to the Trilinos implementation which only
%communicates a volume of data proportional to the number of vertices,
%and ({\it ii}) out-of-core operations in MR-MPI were performed because
%of the page-size restriction, while all Trilinos operations were
%performed in-core.  
%

>>>>

A more detailed analysis of the performance of the MR-MPI PageRank 
algorithm is shown
below.  We perform the analysis using RMAT-20, which is small enough to
fit in memory and thus does not need out-of-core operations, and RMAT-28,
which requires out-of-core operation in MR-MPI.  Table~\ref{table:prdetail}
shows the amount of time spent in each step of Figure~\ref{fig:pr2} for 
the MR-MPI implementation and the Trilinos implementation.  We see that for
both implementation, the matrix-vector multiplication (Step 2) is the most
costly step.  However, the MR-MPI algorithm is consistently slower than
the Trilinos implementation in each step.  This effect is more pronounced
for RMAT-28, where MR-MPI performs out-of-core operations.

\begin{table}[htb]
\begin{center}
\begin{tabular}{|l|r|r|r|r|}
\hline 
& \multicolumn{2}{|c|}{RMAT-20} & \multicolumn{2}{|c|}{RMAT-28} \\
& MR-MPI & Trilinos & MR-MPI & Trilinos  \\
\hline
Step 1 (Compute adjustment) & 0.0203 & 0.00030 &
&  \\
Step 2 (Multiply $A^T x$) & 0.1377 & 0.00455 &
&  \\
Step 3 (Apply adjustment) & 0.0015 & 0.00015 &
&  \\
Step 4 (Scale $y$)  & 0.0013 & 0.00014 &
& \\
Step 5 (Compute residual)  & 0.0207 & 0.00014 &
& \\
\hline
\end{tabular}
\caption{Detailed execution times (in seconds) per PageRank iteration
for each stage of the algorithm
in Figure~\ref{fig:pr2} using MR-MPI and Trilinos.}
\label{table:prdetail}
\end{center}
\end{table}

To understand where MR-MPI spends its effort during the PageRank computation,
we break down the operation times further for the most expensive 
step of the PageRank computation, Step 2; the results are in 
Table~\ref{table:prdetail2}.
In the table, we divide the collate operation into its
two parts:  an aggregate that involves parallel communication and a convert
that gathers local key-value pairs with matching keys into a key-multivalue
pair.  All interprocessor communication occurs in the aggregate operation.
While the time spent in aggregate is significant, it is small compared to
the time spent in the two convert operations.  Further experimentation
is needed to determine whether reorganizing the MapReduce data 
(as described in Section~\ref{subsec:graph_pagerank}) would reduce the
execution time of Step 2.  Organizing the data such that both the matrix
$A$ and vectors $x$ and $y$ are in the same MapReduce object would eliminate
the copy and add functions, as well as the first convert.  However, to update
the $y$ values at the completion of Step 2, either an additional convert
or an aggregate with greater amounts of data would be needed.

STILL NEED RMAT-28 RESULTS AND DISCUSSION OF I/O EFFECT.

\begin{table}[htb]
\begin{center}
\begin{tabular}{|l|r|r|}
\hline 
PageRank Step 2 & RMAT-20 & RMAT-28 \\
\hline
Copy $x$ to $y$ & 1.1\% &  \\
Add $A$ to $y$ & 3.0\% &  \\
Convert $y$ & 23.5\% &  \\
Reduce $y$ & 7.7\% & \\
Aggregate $y$ (first part of collate) & 24.9\% & \\
Convert $y$ (second part of collate) & 38.2\% & \\
Reduce $y$ & 1.6\% & \\
\hline
\end{tabular}
\caption{Percentage of execution time in each operation of Step 2 of the MR-MPI 
PageRank algorithm in Figure~\ref{fig:pr2}.}
\label{table:prdetail2}
\end{center}
\end{table}

Similarly, in Table~\ref{table:prdetail5}, we show the execution times 
required for each operation in Step 5 of Figure~\ref{fig:pr2}, the 
residual computation.  Clearly the time to convert key-value pairs to 
key-multivalues dominates this computation.  Reorganization of the vectors
$x$ and $y$ into a single MapReduce object (as described in 
Section~\ref{subsec:graph_pagerank}) would benefit this step of the computation,
as the add, convert, and reduce operations could be replaced by a single 
map operation, with execution time comparable to the vector-scaling operation
in Step 4.

\begin{table}[htb]
\begin{center}
\begin{tabular}{|l|r|r|}
\hline 
PageRank Step 5 & RMAT-20 & RMAT-28 \\
\hline
Add $y$ to $x$ & 2.2\% &  \\
Convert $x$ & 95.1\% &  \\
Reduce $x$ & 2.5\% & \\
MPI\_Allreduce & 0.2\% & \\
\hline
\end{tabular}
\caption{Percentage of execution time in each operation of Step 5 of the MR-MPI 
PageRank algorithm in  Figure~\ref{fig:pr2}.}
\label{table:prdetail5}
\end{center}
\end{table}
%This result is
%due to two factors: ({\it i}) a higher volume of communication in the
%MapReduce implementation (which comunicates one datum per graph edge
%in each iteration), compared to the Trilinos implementation which only
%communicates a volume of data proportional to the number of vertices,
%and ({\it ii}) out-of-core operations in MR-MPI were performed because
%of the page-size restriction, while all Trilinos operations were
%performed in-core.  
%
